Avoiding global deforestation by taxing land in agricultural production: The implications for global markets

### 2024-03-09 paper.Rmd
Update paper.Rmd

---
title: "Avoiding global deforestation by taxing land in agricultural production: The implications for global markets"
date: "`r format(Sys.time(), '%B %d, %Y')`"
author: "Eric C. Davis, Maros Ivanic, and Brent Sohngen"
output: 
  bookdown::pdf_document2:
    keep_tex: TRUE
toc: FALSE
bibliography: references.bib
header-includes:
    - \usepackage{setspace}\doublespacing
abstract: "The projected growth in population and incomes is expected to create pressure to convert forestland into farmland. At the same time, the increasingly negative climate impacts are expected to generate further pressure to enhance the terrestrial carbon sink. Even though these goals are  incompatible as reversing the deforestation trend by afforesting cropland would result in negative market impacts such as higher food prices, using the GTAP and GTM models, we find that these impacts would be relatively small if the goal of  preserving 144.2 million hectares of forestland that otherwise would be converted to agricultural land by 2033 is achieved through a tax on using land in agricultural production. As to the economic price for doing so, the avoided deforestation would in most regions of the world result in less agricultural output and higher market prices. This is estimated to impact the well-being of global consumers by $92.1 billion, which translates to a global average cost of $10.6 per person in 2033."
---

@@

Both the GTAP and GTM models represent widely used models for policy analysis. In the case of GTAP, the model is used to describe the implications of various policy instruments for national and global markets, including prices, output and trade. The GTM model is, on the other hand, capable to translating the impacts of population and global growth, and the changes in land returns, to changes in global forest stocks and calculate the impacts on total carbon contained in forests.

>>>>

---
title: "Avoiding global deforestation by taxing land in agricultural production: The implications for global markets"
date: "`r format(Sys.time(), '%B %d, %Y')`"
author: "Eric C. Davis, Maros Ivanic, and Brent Sohngen"
output: 
  bookdown::pdf_document2:
    keep_tex: TRUE
toc: FALSE
bibliography: references.bib
header-includes:
    - \usepackage{setspace}\doublespacing
abstract: "The projected global growth in population and incomes is expected to create pressure to convert forestland into farmland, with an estimated 144.2 million hectares of forestland estimated to be converted to agricultural land by 2033. At the same time, increasingly negative climate impacts are expected to generate further pressure to enhance the terrestrial carbon sink, but taking such actions would likely result in less agricultural output and higher food prices. Using the GTAP and GTM models, we model the impact of applying a tax on converting forestland for use in agricultural production and find that the net deforestation rate could be brought to zero for a relatively small cost. Results suggest that the well-being of global consumers would be reduced by $92.1 billion in 2033, which translates to a global average annual cost of $10.6 per person."
---

@@

Both the GTAP and GTM models represent widely used models for policy analysis. In the case of GTAP, the model is used to describe the implications of various policy instruments for national and global markets, including prices, output, and trade. The GTM model is, on the other hand, capable to translating the impacts of population and global growth, and the changes in land returns, to changes in global forest stocks and calculate the impacts on total carbon contained in forests.
